systemfit: A Package for Estimating\\ Systems of Simultaneous Equations in \proglang{R}

### 2007-06-06 paper/trunk/systemfit_estimation.tex
a few minor improvements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@712 591db2bd-e0a0-4d22-8c19-2192dbb7da31

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
   \label{eq:ols}
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
   \label{eq:olsCovSameSigma}
\end{equation}
where $\sigma^2 = \E \left[ u_{it}^2 \right]$.
This assumes that the disturbances of all equations have the
same variance.

@@

The standard 3SLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
   \label{eq:3slsGls}
\end{equation}
where $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
Its covariance matrix is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} .
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] .
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \Omega^{-1} X
   \right)^{-1} \widehat{X}' \Omega^{-1} y .
   \label{eq:3slsIv}
\end{equation}
The covariance matrix of this 3SLS-IV estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the Generalized Method of Moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y .
   \label{eq:3slsGmm}
\end{equation}
The covariance matrix of the 3SLS-GMM estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-GMM} \right] =
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \Omega^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y .
   \label{eq:3slsSchmidt}
\end{equation}
The covariance matrix of this estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} .
\end{equation}
The econometrics software EViews uses
\begin{equation}
   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
as defined by (\ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate the covariance matrix of the 3SLS estimator.

>>>>

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
   \label{eq:ols}
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
   \label{eq:olsCovSameSigma}
\end{equation}
where $\sigma^2 = \E \left[ u_{it}^2 \right] \, \forall \, i$.
This assumes that the disturbances of all equations have the
same variance.

@@

The standard 3SLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
   \label{eq:3slsGls}
\end{equation}
where $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
Its covariance matrix is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} .
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] .
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \Omega^{-1} X
   \right)^{-1} \widehat{X}' \Omega^{-1} y .
   \label{eq:3slsIv}
\end{equation}
The covariance matrix of this 3SLS-IV estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the Generalized Method of Moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y .
   \label{eq:3slsGmm}
\end{equation}
The covariance matrix of the 3SLS-GMM estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-GMM} \right] =
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \Omega^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y .
   \label{eq:3slsSchmidt}
\end{equation}
The covariance matrix of this estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} .
\end{equation}
The econometrics software \proglang{EViews} uses
\begin{equation}
   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
as defined by (\ref{eq:beta2sls}).
\proglang{EViews} uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate the covariance matrix of the 3SLS estimator.
